---
title: "- Missing Data"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r}

#We are using a .dat file (survey.dat) created from the SPSS file survey.sav  taken from SPSS Survival Manual 6th Edition Julie Pallant
#http://spss.allenandunwin.com.s3-website-ap-southeast-2.amazonaws.com/data-files.html#.Wb0vvnWP-po
#Results on a survey on well being
#We need to load the file so that we can use it in R.
sdata <- read.table("survey.dat")

#Setting the column names to be that used in the dataset
colnames(sdata) <- tolower(colnames(sdata))

needed_packages <- c("VIM", "tidyverse", "naniar")                      
# Extract not installed packages
not_installed <- needed_packages[!(needed_packages %in% installed.packages()[ , "Package"])]    
# Install not installed packages
if(length(not_installed)) install.packages(not_installed) 
library("VIM")
library("tidyverse")
library("naniar")


```
## Visualise the missing data level and pattern
```{r}

#Our variables of interest are the nine continuous variables
#We create a subset with just these variables

varsint<-c("toptim",     "tmast", "tposaff", "tnegaff", "tlifesat", "tpstress","tslfest", "tmarlow" , "tpcoiss" )
sdatasubset<-sdata[varsint]#create a subset of data with just the variables of interest
summary(sdatasubset)#Get our summary statistics - these are just for us not for publication

#We want to look at levels of missingness using the aggr function from the VIM package
#We store the results in a variable called res 

res<-summary(VIM::aggr(sdatasubset, sortVar=TRUE))$combinations

#Output the most common combinations
head(res[rev(order(res[,2])),])
 

#We cab create a plot for our 9 variables of interest showing the percentage missing using vis_miss from the naniar package
naniar::vis_miss(sdatasubset)


#To look at the number of cases and the combinations of missingness you can use gg_miss_upset

naniar::gg_miss_upset(sdatasubset)

#tpcoiss alone is missing - 5 cases
#tpstress is missing - 3 cases
#tmarlow is missing - 2 cases
#Then 1 case where all variables are missing
#There are 2 cases where tnegaff, tpstress and toptim are missing 
#etc.
#*** There is no real pattern to this missing data

#To look at patterns by group you need to include the grouping value as part of the sub-set
#You can then look at the pattern using gg_mis_fct from naniar package

varsint<-c("toptim",     "tmast", "tposaff", "tnegaff", "tlifesat", "tpstress","tslfest", "tmarlow" , "tpcoiss","smoke" )

sdatasubset<-sdata[varsint]#create a subset of data with just the variables of interest

naniar::gg_miss_fct(x = sdatasubset, fct = smoke)

## You can create tables to look at patterns of missing data in continuous for nominal
## Or use some of the packages available in  https://www.tidyverse.org/

