---
title: "Difference - Anova Survey"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---


## SETUP
```{r setup}

#We are using a .dat file (survey.dat) created from the SPSS file survey.sav  taken from SPSS Survival Manual 6th Edition Julie Pallant
#http://spss.allenandunwin.com.s3-website-ap-southeast-2.amazonaws.com/data-files.html#.Wb0vvnWP-po
#Results on a survey on well being
#We need to load the file so that we can use it in R.


#We need to load the file so that we can use it in R.
survey <- read.table("survey.dat")
#Setting the column names to be that used in the dataset
colnames(survey) <- tolower(colnames(survey))

needed_packages <- c("pastecs", "ggplot2", "psych", "semTools", "FSA",  "rstatix", "effectsize")                      
# Extract not installed packages
not_installed <- needed_packages[!(needed_packages %in% installed.packages()[ , "Package"])]    
# Install not installed packages
if(length(not_installed)) install.packages(not_installed) 

library(pastecs) #For creating descriptive statistic summaries
library(ggplot2) #For creating histograms with more detail than plot
library(psych) # Some useful descriptive functions
library(semTools) #For skewness and kurtosis
library(FSA) #For percentage
library(rstatix) #To calculate post-hoc test
library(effectsize)
```

## 1. H0: There are no differences between total optimism scores for respondents in Group 1, Group 2 and Group 3. HA: There are a differences between  total optimism scores for repondents in Group 1, Group 2 and Group 3.
### Step 1: Examine/Assess Normality
```{r toptim-normality}

#We will allocate the histogram to a variable to allow use to manipulate it
gg <- ggplot(survey, aes(x=toptim))
gg <- gg+ggtitle("Figure 1 - Histogram for Total Optimism Scores")

#Change the label of the x axis
gg <- gg + labs(x="Total Optimism Scores")

#manage bin width and colours
gg <- gg + geom_histogram(binwidth=2, colour="black", aes(y=..density.., fill=..count..))
gg <- gg + scale_fill_gradient("Count", low="#DCDCDC", high="#7C7C7C")

#adding a normal curve
#use stat_function to compute a normalized score for each value of toptim
#pass the mean and standard deviation
#use the na.rm parameter to say how missing values are handled
gg <- gg + stat_function(fun=dnorm, color="red",args=list(mean=mean(survey$toptim, na.rm=TRUE), sd=sd(survey$toptim, na.rm=TRUE)))

#to display the graph request the contents of the variable be shown
gg

#Create a qqplot
qqnorm(survey$toptim, main="Figure 2 - QQ Plot for Total Optimism")
qqline(survey$toptim, col=2) #show a line on the plot


#stat.desc is a function from pastecs - make sure you include the basic switch=F to ensure you don't get scientific notation
pastecs::stat.desc(survey$toptim, basic=F)

#We can make our decision based on the value of the standardized score for skew and kurtosis
#We divide the skew statistic by the standard error to get the standardized score
#This will indicate if we have a problem
tpskew<-semTools::skew(survey$toptim)
tpkurt<-semTools::kurtosis(survey$toptim)
tpskew[1]/tpskew[2]

tpkurt[1]/tpkurt[2]

#and by calculating the percentage of standardized scores for the variable itself that are outside our acceptable range
#This will tell us how big a problem we have
# Calculate the percentage of standardized scores that are greater than 1.96
# the perc function which is part of the FSA package which calculate the percentage that are within a range - you can look for greater than "gt", greater than or equal "geq", "gt", less than or equal "leq",  or less than "lt"),
# scale is a function that creates z scores, abs gets absolute value

ztoptim<- abs(scale(survey$toptim))

FSA::perc(as.numeric(ztoptim), 1.96, "gt")
FSA::perc(as.numeric(ztoptim), 3.29, "gt")

```
### Report assessment of normality

<p style="text-align:justify">
Total optimism scores data was assessed for normality. Visual inspection of the histogram and QQ-Plot (see Figure 1 and Figure 2) identified some issues with skewness and kurtosis. The standardized score for kurtosis (`r round(tpkurt[1]/tpkurt[2],2)`) can be considered acceptable using the criteria proposed by West, Finch and Curran (1996), but the standardized score for skewness (`r round(tpskew[1]/tpskew[2],2)`)  was outside the acceptable range.  However over 99% of standardized scores for total optimism and associated sensation fall within the bounds of +/- 3.29, using the guidance of Field, Miles and Field (2013) the data can be considered to approximate a normal distribution (m=`r round(mean(survey$toptim, na.rm=TRUE),2)`, sd=`r round(sd(survey$toptim, na.rm=TRUE),2)`, n=`r length(survey$toptim)-sum(is.na(survey$toptim))`).
</p>

### Differences more than 2 groups - Parametric Tests
#### ANOVA
##### Describe Total Optimism by group (Group 1, Group 2, Group3 agegp3)

```{r Describe}
#Get descriptive statistics by group - output as a matrix
psych::describeBy(survey$toptim, survey$agegp3, mat=TRUE)
#Store the output to use in our final reporting of the outcomes of ANOVA

optdesc<-psych::describeBy(survey$toptim,survey$agegp3,mat=TRUE)

#Conduct Bartlett's test for homogeneity of variance in library car - the null hypothesis is that variances in groups are equal so to assume homogeneity we would expect probability to not be statistically significant.
stats::bartlett.test(toptim~ agegp3, data=survey)
#p value is > 0.05 so the result is not statistically significant so we can assume homogeneity


#Conduct ANOVA using the one-way ANOVA from the stats package

# Compute the analysis of variance using the oneway.test from the stats package- we set VAR.equal to be true because we have homogeneity
res <- stats::oneway.test(toptim ~ as.factor(agegp3), data = survey, var.equal = TRUE)
# Summary of the analysis
res

#We use Tukey for post-hoc testing
rstatix::tukey_hsd(survey, toptim ~ agegp3)

#Compute our Eta squared
effes=effectsize::effectsize(res)
effes

#Store the relevant pieces of the output from ANOVA in variables to use for reporting
#Degrees of freedom
df1=res$parameter[1]
df2=res$parameter[2]
#F statistic
Fstat=round(res$statistic, 3)
#Pvalue
pval=round(res$p.value,2)


#A statistically significant difference was found so we need to examine and report the post-hoc results also.
```


### Reporting the results with eta squared effect

<p>



A one-way between-groups analysis of variance (ANOVA) was conducted to explore the impact of age on levels of optimism, as measured by the Life orientation Test (LOT). Participants were divided into three groups according to their age (Group 1: 29 yrs or less; Group 2: 30 to 44 yrs; Group 3: 45yrs and above).  There was a statistically significant difference at the p < .05 level in LOT scores for the three age groups: (F(`r df1`, `r df2`)= `r Fstat`, p<0.05. 
Despite reaching statistical significance, the actual difference in mean scores between groups was quite small. The effect size, calculated using eta squared was (`r round(effes$Eta2,2)`). 
Post-hoc comparisons using the Tukey HSD test indicated that the mean score for Group 1 
(M=`r round(optdesc$mean[1],2)`,SD=`r round(optdesc$sd[1],2)`) 
was significantly different to that for  Group 3  (M=`r round(optdesc$mean[3],2)`,SD=`r round(optdesc$sd[3],2)`). Group 2 (M=`r round(optdesc$mean[2],2)`,SD=`r round(optdesc$sd[2],2)`) did not differ significantly from Group 1 or Group3.
</p>



