```{r}
#install prerequisite packages, and set workd dir
library("rstudioapi")
library(haven)
library(ggplot2)

setwd(dirname(getActiveDocumentContext()$path))
install_packages <- function(pkg) { 
  
  # Install package if it is not already
  if (!(pkg %in% installed.packages()[, "Package"])){ 
    
    install.packages(pkg, repos='http://cran.us.r-project.org')
  }
  
  library(pkg, character.only = TRUE)
  
} # end installPackages()

#Create the list of packages we need
pkg_list = c("tidyverse", "ggpubr")
#Call our function passing it the list of packages
lapply(pkg_list, install_packages)

```

```{r}
survey <- read_sav("survey.sav")

#summary(survey)
summary(survey$tposaff)
```

```{r}
summary(survey$tpstress)
```

## Step 2: Try to write the code to investigate the following relationships using a Pearson correlation test

#### Relationship between Total Positive Affect and Total Perceived Stress

```{r}
hist_tposaff <- ggplot(survey, aes(x=tposaff)) + geom_histogram(fill="black",binwidth = 0.4)
hist_tpstress <- ggplot(survey, aes(x=tpstress)) + geom_histogram(fill="black",binwidth = 0.4)
ggarrange(hist_tposaff, hist_tpstress, 
          ncol = 2)
```
  
The data from both variables looks from a normal distribution but is it?
```{r}
shapiro.test(survey$tposaff)
```

```{r}
shapiro.test(survey$tpstress)
```
The data doesn't seem to belong to a normal distribution, which correlation method should I use?

```{r}
ggplot(survey, aes(x=tposaff, y=tpstress)) + geom_point() + geom_smooth(method=lm)
```

```{r}
cor.test(survey$tposaff, survey$tpstress, method = "pearson")
```
```{r}
cor.test(survey$tposaff, survey$tpstress, method = "kendall")
```

```{r}
cor.test(survey$tposaff, survey$tpstress, method = "spearman")
```


#### Relationship between Total Life Satisfaction and Total Perceived Control of Internal States.

```{r}
ggplot(survey, aes(x=tlifesat, y=tpcoiss)) + geom_point() + geom_smooth(method=lm)

```

```{r}
cor.test(survey$tlifesat, survey$tpcoiss, method = "pearson")
```

####	Relationship between Total Mastery and Total Optimism.

```{r}
ggplot(survey, aes(x=tmast, y=toptim)) + geom_point() + geom_smooth(method=lm)
```


```{r}
cor.test(survey$tmast, survey$toptim, method = "pearson")
```

```{r}
cor.test(survey$tmast, survey$toptim, method = "kendall")
```

```{r}
cor.test(survey$tmast, survey$toptim, method = "spearman")
```
